Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-24 cav2017/cav2017.tex
add pseudocode for stability, some formatting to Alessandro's proofs

%\todo{please eventually itemize next}
\begin{enumerate}
\item The inductive synthesis phase (i.e. {\sc synthesise}) uses BMC to
compute a candidate solution $K$
that satisfies both the stability criteria detailed in 
Sec.~\ref{ssec:stability} and the safety specification in Sec.~\ref{ssec:safety}.
%% The stability criteria are satisfied if the characteristic polynomial of $A_d - B_d * K_d$ meets Jury's criteria. This is 
%% independent of the initial states. 
Regarding the latter, we synthesise a safe controller by unwinding the transition system
$k$ steps and picking a controller $K$ and an initial state such that the states at each step do
not violate the safety criteria. This is sound if the current $k$ is below the
completeness threshold, which we assume to be true in the {\sc synthesise} phase.
We also assume some
precision $\langle I_p,F_p\rangle$ for the plant and a sampling rate.  The
checks that all these assumptions are sound are performed by subsequent
{\sc verify} stages.

\begin{algorithm}[]
\begin{algorithmic}[1]
\Function{$safetyCheck()$}{}
	\State assert($ \underline{u}  \leq u \leq \overline{u}$)
	\While{$i=0; i< k; i++$}
		%\State $u = (\langle I_p,F_p\rangle)((\langle I_c,F_c\rangle)K * (\langle I_c,F_c\rangle) x)$
		\State $u = (plant\_typet)((controller\_typet)K * (controller\_typet) x)$
		\State $x = A * x + B * u$
		\State assert($\underline{x} \leq x \leq \overline{x}$ )
  	\EndWhile
  %\State Return $K$
\EndFunction
\end{algorithmic}
\label{alg:safetycheck}
\end{algorithm}

@@

\todo[inline]{please will Alessandro write out the proof that checking only the corner cases is valid for the safety verification - EP}

Proof. 
Consider the set of initial states, $X_0$, assumed to be convex since it's a box. 
Name $v_i$ its vertexes, where $i=1,\ldots, 2^n$.  
Thus any point $x \in X_0$ can be expressed by convexity as $x = \sum_{i=1}^{2^n} \alpha_i v_i$, 
where $\sum_{i=1}^{2^n} \alpha_i =1$. 
So then if $x_0=x$, then $x_k = (A_d - B_d K)^k x = 
(A_d - B_d K)^k \sum_{i=1}^{2^n} \alpha_i v_i = 
\sum_{i=1}^{2^n} \alpha_i (A_d - B_d K)^k v_i  = 
\sum_{i=1}^{2^n} \alpha_i x_k^i$, 
where I have denoted with $x_k^i$ the trajectories obtained from the single vertexes $v_i$.QED

@@

Proof sketch. 
Consider spectrum of matrix $(A_d - B_d K)$ and select eigenvalue with smallest (non trivial) imaginary value.  
call such value $\vartheta$. 
At every time interval $kT_s, (k+1)T_s$ ($T_s$ being the sampling time), 
the dynamics projected on the corresponding eigenspace rotate of $\vartheta T_s$ radiants.  
So surely taking $k$ as the ceiling of $\frac{2\pi}{\vartheta T_s}$, after $k$ steps we've completed a rotation. 
This is the CT.

\item The last {\sc verify} stage, {\sc sampling}, 
ensures that the current sampling rate is large enough by using an 
external checker (\addtodo{add more details here once we have them}).

>>>>

%\todo{please eventually itemize next}
\begin{enumerate}
\item The inductive synthesis phase (i.e. {\sc synthesise}) uses BMC to
compute a candidate solution $K$
that satisfies both the stability criteria detailed in 
Sec.~\ref{ssec:stability} and the safety specification in Sec.~\ref{ssec:safety}.
To synthesise a controller that satisfies the stability criteria, we require that
a computed characteristic polynomial satisifies Jury's criteria. The details of this
calculation can be found in the Appendix.

%% The stability criteria are satisfied if the characteristic polynomial of $A_d - B_d * K_d$ meets Jury's criteria. This is 
%% independent of the initial states. 
Regarding the latter, we synthesise a safe controller by unwinding the transition system
$k$ steps and picking a controller $K$ and an initial state such that the states at each step do
not violate the safety criteria. This is sound if the current $k$ is below the
completeness threshold, which we assume to be true in the {\sc synthesise} phase.
We also assume some
precision $\langle I_p,F_p\rangle$ for the plant and a sampling rate.  The
checks that all these assumptions are sound are performed by subsequent
{\sc verify} stages.

\begin{algorithm}[]
\begin{algorithmic}[1]
\Function{$stabilityCheck()$}{}
  \State computecharPoly($A_d - B_dK$)
  \State assert(Jury's criteria hold)
  %\State Return $K$
\EndFunction
\end{algorithmic}
\label{alg:stabilitycheck}
\end{algorithm}

\begin{algorithm}[]
\begin{algorithmic}[1]
\Function{$safetyCheck()$}{}
	\State assert($ \underline{u}  \leq u \leq \overline{u}$)
  \State assert($ \underline{x_0}  \leq x_0 \leq \overline{x_0}$)
	\While{$i=0; i< k; i++$}
		%\State $u = (\langle I_p,F_p\rangle)((\langle I_c,F_c\rangle)K * (\langle I_c,F_c\rangle) x)$
		\State $u = (plant\_typet)((controller\_typet)K * (controller\_typet) x)$
		\State $x = A * x + B * u$
		\State assert($\underline{x} \leq x \leq \overline{x}$ )
  	\EndWhile
  %\State Return $K$
\EndFunction
\end{algorithmic}
\label{alg:safetycheck}
\end{algorithm}

@@

\begin{proof} 
Consider the set of initial states, $X_0$, assumed to be convex since it's a hypercube. 
Name $v_i$ its vertexes, where $i=1,\ldots, 2^n$.  
Thus any point $x \in X_0$ can be expressed by convexity as $x = \sum_{i=1}^{2^n} \alpha_i v_i$, 
where $\sum_{i=1}^{2^n} \alpha_i =1$. \\
So then if $x_0=x$, then 
\begin{align*}
x_k   &= (A_d - B_d K)^k x \\
      &= (A_d - B_d K)^k \sum_{i=1}^{2^n} \alpha_i v_i \\
      &= \sum_{i=1}^{2^n} \alpha_i (A_d - B_d K)^k v_i \\
      &= \sum_{i=1}^{2^n} \alpha_i x_k^i
 \end{align*}     
where $x_k^i$ denotes the trajectories obtained from the single vertexes $v_i$. 
\end{proof}

@@

\begin{proof}
Consider spectrum of matrix $(A_d - B_d K)$ and select eigenvalue with smallest (non trivial) imaginary value.  
call such value $\vartheta$. 
At every time interval $kT_s, (k+1)T_s$ ($T_s$ being the sampling time), 
the dynamics projected on the corresponding eigenspace rotate of $\vartheta T_s$ radiants.  
So surely taking $k$ as the ceiling of $\frac{2\pi}{\vartheta T_s}$, after $k$ steps we've completed a rotation. 
This is the completeness threshold. 
\end{proof}
\item The last {\sc verify} stage, {\sc sampling}, 
ensures that the current sampling rate is large enough by using an 
external checker (\addtodo{add more details here once we have them}).
